Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering

### 2014-09-11 evaluation.tex
Update Evaluation.

In order to explore the generality of the architecture for
different monitors, we implemented three different monitors: uninitialized
memory check (UMC), array bounds check (BC), and dynamic information flow
tracking (DIFT).  Uninitialized memory check seeks to detect loading from
memory locations that are not initialized first.  Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from.

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC and DIFT
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 1 billion instructions. An initial slack of 1
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets.

@@

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 4x to 11x with an average
of 7x. BC shows normalized execution times of 9-28x with an average of
18x while DIFT shows normalized execution times of 6x-17x with an average of
11x. One of the reasons for these high overheads is that our implementations of
these monitors dynamically allocate memory for new metadata. This can be an
expensive, many-cycle operation. Statically allocating memory for metadata
will reduce the execution time but requires a large upfront memory footprint
which caused issues with our simulator.

@@

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 2-6x with an average of 4x with null metadata filtering.
This is an average reduction in overheads of 52\%. For BC, normalized execution
times drop from 18x down to 4x on average (an 82\% reduction). The range of normalized execution
times for filtered BC is 1.1-17x. This is not surprising since in the baseline
implementation all loads and stores needed to be monitored. However, with
filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
on average after filtering, a 90\% reduction in monitoring overheads, with a
maximum overhead of 42\% after filtering. This is due, in part, to the fact
that
for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering.

@@

% BC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
    \vspace{-0.4in}
    \caption{Coverage versus varying overhead budget for array bounds check.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure*}

% UMC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.4in}
    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Although the overheads of DIFT are quite low after filtering, BC and UMC still
show significant overheads. In this section, we evaluate the effectiveness of
using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system. 
While we could not evaluate real errors/attacks due to difficulty in setting up
a large number of bugs and exploits, we believe that the percentage of checks
provides a good estimate of detection probability when errors/attacks are 
uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average.
By increasing this overhead budget to 50\%, UMC achieves 51\% coverage. Even
higher coverage can be achieved by allowing higher overheads.

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC at 10\% overhead with an unrestricted dropping policy.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by introducing some randomness into the
dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running with a 10\%
overhead budget and an unrestricted dropping policy. Each run was simulated for
100 million instructions. For these simulations, events were
dropped probabilistically with the probability scaled proportionally to slack.
That is, when slack is high, the probability of dropping is low and the
probability increases as we run out of slack. We see that the coverage achieved
quickly asymptotes for an unrestricted dropping policy. This is due to the fact
that since there are multiple dependent monitoring operations that must be done
in order for a final check to be successful, a small amount of randomness at
each event quickly compounds to reduce the probability of certain checks to
near zero. Instead, source-only dropping is much better suited for achieving high
coverage over multiple runs.

@@

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
less than 1\% of the main core's peak power usage. Similarly, the average runtime power is 3-4 
mW, corresponding to about 1\% of the main core's runtime power.

>>>>

In order to explore the generality of the architecture for
different monitors, we implemented three different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), and instruction-mix profiling (IMP).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations. Additionally, recent
work has shown that this information can be used to detect malicious software \cite{tang-raid14}.

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 2 billion instructions. An initial slack of 2
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets.

@@

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads for UMC, BC, DIFT, and IMP.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 5x to 16x with an average
of 10x. BC shows normalized execution times of 8-24x with an average of
16x while DIFT shows normalized execution times of 4x-14x with an average of
9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
These monitors show high overheads because they require several
instructions to process each monitored instruction from the main core. Even
small increases in the number of instructions needed by the monitor lead to
high overheads when multiplied across the main program's execution. For
example, because of the large metadata requirements of UMC, BC, and DIFT, these
monitors must dynamically allocate memory for new metadata. On the other hand,
IMP does not require this dynamic allocation and is overall a simple monitor.
Thus, its overheads are much lower.

@@

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 1.7-7x with an average of 4x with null metadata filtering.
This is an average reduction in overheads of 65\%. For BC, normalized execution
times drop from 16x down to FIXMEx on average (an FIXME\% reduction). The range of normalized execution
times for filtered BC is FIXME-FIXMEx. This is not surprising since in the baseline
implementation all loads and stores needed to be monitored. However, with
filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
on average after filtering, an 88\% reduction in monitoring overheads, with a
maximum overhead of 31\% after filtering. This is due, in part, to the fact
that
for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering. IMP does not have
null metadata and thus is not affected by null metadata filtering.

@@

% BC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for array bounds check.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure*}

% UMC sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Error versus varying overhead budget for instruction-mix profiling.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Although the overheads of DIFT are quite low after filtering, BC and UMC still
show significant overheads. Similarly, the overheads of IMP are not improved by filtering. In this section, we evaluate the effectiveness of
using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system. 
While we could not evaluate real errors/attacks due to difficulty in setting up
a large number of bugs and exploits, we believe that the percentage of checks
provides a good estimate of detection probability when errors/attacks are 
uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 10\% overhead budget, UMC achieves 30\% monitoring coverage on average.
By increasing this overhead budget to 50\%, UMC achieves 48\% coverage. Even
higher coverage can be achieved by allowing higher overheads.

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We compare these percentages against the case when full monitoring
is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the max error across
instruction types for each benchmark and for varying overheads. We see that
with a 10\% overhead a max error of 9\% is observed and an average of less than
4\%. This error decreases as overhead is increased.

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by randomly deciding whether events
should be dropped.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies. Here, we use a 10\% probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that since there are multiple dependent
monitoring operations that must be done in order for a final check to be
successful, a small amount of randomness at each event quickly compounds to
reduce the probability of certain checks to near zero. Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.

@@

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 50\% monitoring overhead target. The peak power is 41-72 mW, which is
less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43
mW, corresponding to 5-11\% of the main core's runtime power.
